A Journey from Wild to Textbook Data to Reproducibly Refresh the Wages Data from the National Longitudinal Survey of Youth Database

### 2022-02-16 paper/paper.tex
small fixes all over plus new check code for hgc

% Pandoc citation processing

\usepackage{tcolorbox}
\usepackage{fontawesome}
\usepackage{color, colortbl}
\usepackage{booktabs}
\usepackage{longtable}
\usepackage{array}
\usepackage{multirow}
\usepackage{wrapfig}
\usepackage{float}
\usepackage{colortbl}
\usepackage{pdflscape}
\usepackage{tabu}
\usepackage{threeparttable}
\usepackage{threeparttablex}
\usepackage[normalem]{ulem}
\usepackage{makecell}
\usepackage{xcolor}

\begin{document}
\maketitle

\def\tightlist{}

@@

Our process of cleaning builds heavily on the \texttt{tidyverse} approach (Wickham, Averick, et al. 2019). The data is first organized into ``tidy data'' (Wickham 2014) and then further wrangled using the data pipeline and split-apply-combine approach (Wickham 2011). Tidy data shouldn't be confused with ``tame data'' which Kim, Ismay, and Chunn (2018) coined to refer to textbook data sets suitable for teaching, particularly teaching statistics. The resulting (tame) data is provided in a new R package called \texttt{[CENSORED]} which includes the code so that the process is reproducible, and could be used to further refresh the data as new records are made available in the NLSY79 database.

@@

\begin{itemize}
\tightlist
\item
  \texttt{NLSY79.csv}: Comma Separated Value format of the response data,
\item
  \texttt{NLSY79.dat}: .dat format of the response data,
\item
  \texttt{NLSY79.NLSY79}: Tagset of variables that can be uploaded to the website to recreate the data set, and
\item
  \texttt{NLSY79.R}: R script for reading the data into R and converting the variables' names and label into something more sensible.
\end{itemize}

@@

Thus we re-arrange and wrangle the data into tidy data form, columns corresponding to individual ID, year, job number, wage in dollars and the demographic variables. This is done using the \texttt{tidyverse} suite of packages (Wickham, Averick, et al. 2019): \texttt{tidyr} (Wickham 2020) to pivot the data into long form, with \texttt{dplyr} (Wickham, François, et al. 2020) and \texttt{stringr} (Wickham 2019) to mutate new variables from the downloaded data from the database, and code levels of factors by text wrangling. The long form of the data makes it possible to do these data transformations efficiently, and it is an intermediate step towards the final target data. The code for tidying the data are demonstrated at \url{CENSORED/articles/raw-to-input-data.html} but also described in the subsequent subsections.

@@

{\centering \includegraphics[width=0.85\linewidth]{/Users/deeamaliah/Documents/yowie-paper/paper/figures/flowchart_blind}

@@

\hypertarget{refs}{}
\leavevmode\hypertarget{ref-iris-data}{}%
Andereson, Edgar. 1935. ``The Irises of the Gaspe Peninsula.'' \emph{Bulletin of the American Iris Society} 59: 2--5.

\leavevmode\hypertarget{ref-nlsy79}{}%
Bureau of Labor Statistics, U.S. Department of Labor. 2021a. ``National Longitudinal Survey of Youth 1979 Cohort, 1979-2016 (Rounds 1-28).'' Produced and distributed by the Center for Human Resource Research (CHRR), The Ohio State University. Columbus, OH, through \url{https://www.nlsinfo.org/bibliography-citing-nls-data}.

\leavevmode\hypertarget{ref-nlsy79guide}{}%
---------. 2021b. ``National Longitudinal Survey of Youth 1979 Cohort, Topical Guide to the Data.'' \url{https://www.nlsinfo.org/content/cohorts/nlsy79/topical-guide/employment/work-experience}.

\leavevmode\hypertarget{ref-shiny}{}%
Chang, Winston, Joe Cheng, JJ Allaire, Yihui Xie, and Jonathan McPherson. 2020. \emph{shiny: Web Application Framework for R}. \url{https://CRAN.R-project.org/package=shiny}.

\leavevmode\hypertarget{ref-Chatfield1985TIEo}{}%
Chatfield, C. 1985. ``The Initial Examination of Data.'' \emph{Journal of the Royal Statistical Society. Series A. General} 148 (3): 214--53.

\leavevmode\hypertarget{ref-eliznlsy}{}%
Cooksey, Elizabeth C. 2017. ``Using the National Longitudinal Surveys of Youth (Nlsy) to Conduct Life Course Analyses.'' In \emph{Handbook of Life Course Health Development}, edited by Richard M. Lerner Neal Halfon Christoper B. Forrest, 561--77. Cham: Springer. \url{https://doi.org/https://doi.org/10.1007/978-3-319-47143-3_23}.

\leavevmode\hypertarget{ref-DasuTamraparni2003Edma}{}%
Dasu, Tamraparni, and Theodore Johnson. 2003. \emph{Exploratory Data Mining and Data Cleaning}. Wiley Series in Probability and Statistics. Hoboken: WILEY.

\leavevmode\hypertarget{ref-janitor}{}%
Firke, Sam. 2020. \emph{janitor: Simple Tools for Examining and Cleaning Dirty Data}. \url{https://CRAN.R-project.org/package=janitor}.

\leavevmode\hypertarget{ref-racismnotrace}{}%
Fullilove, M. T. 1998. ``Comment: Abandoning ``Race" as a Variable in Public Health Research--an Idea Whose Time Has Come.'' \emph{American Journal of Public Health} 88 (9): 1297--8.

\leavevmode\hypertarget{ref-grimshaw}{}%
Grimshaw, Scott D. 2015. ``A Framework for Infusing Authentic Data Experiences Within Statistics Courses.'' \emph{The American Statistician} 69 (4): 307--14. \url{https://doi.org/10.1080/00031305.2015.1081106}.

\leavevmode\hypertarget{ref-purrr}{}%
Henry, Lionel, and Hadley Wickham. 2020. \emph{purrr: Functional Programming Tools}. \url{https://CRAN.R-project.org/package=purrr}.

\leavevmode\hypertarget{ref-penguins-data}{}%
Horst, Allison Marie, Alison Presmanes Hill, and Kristen B Gorman. 2020. \emph{Palmerpenguins: Palmer Archipelago (Antarctica) Penguin Data}. \url{https://doi.org/10.5281/zenodo.3960218}.

\leavevmode\hypertarget{ref-HuebnerMariannePhD2016Asat}{}%
Huebner, Marianne, Werner Vach, and Saskia le Cessie. 2016. ``A Systematic Approach to Initial Data Analysis Is Good Research Practice.'' \emph{The Journal of Thoracic and Cardiovascular Surgery} 151 (1): 25--27.

\leavevmode\hypertarget{ref-ilk2004}{}%
Ilk, Ozlem. 2004. ``Exploratory Multivariate Longitudinal Data Analysis and Models for Multivariate Longitudinal Binary Data.'' PhD thesis, Iowa State University. \url{https://doi.org/10.31274/rtd-180813-11012}.

\leavevmode\hypertarget{ref-tamedata}{}%
Kim, A. Y, C. Ismay, and J. Chunn. 2018. ``The Fivethirtyeight R Package: ``Tame Data" Principles for Introductory Statistics and Data Science Courses.'' \emph{Technology Innovations in Statistics Education} 11 (1). \url{https://doi.org/10.5070/T511103589}.

\leavevmode\hypertarget{ref-KollerManuel2016rARP}{}%
Koller, Manuel. 2016. ``robustlmm: An R Package for Robust Estimation of Linear Mixed-Effects Models.'' \emph{Journal of Statistical Software} 75 (6): 1--24.

\leavevmode\hypertarget{ref-notaverage}{}%
Moncrief, Marc. 2015. ``By the Numbers - the Average Australian Doesn't Exist ... Not a Single One of Us Is 'Normal'.'' \url{https://bit.ly/smh-not-normal}.

\leavevmode\hypertarget{ref-opendata}{}%
Open Knowledge Foundation. 2021. ``Open Definition. Defining Open in Open Data, Open Content, and Open Knowledge.'' 2021. \url{http://opendefinition.org/od/2.1/en/}.

\leavevmode\hypertarget{ref-patchwork}{}%
Pedersen, Thomas Lin. 2020. \emph{patchwork: The Composer of Plots}. \url{https://CRAN.R-project.org/package=patchwork}.

\leavevmode\hypertarget{ref-MichaelRPergamit2001DWTN}{}%
Pergamit, Michael R., Charles R. Pierret, Donna S. Rothstein, and Jonathan R. Veum. 2001. ``Data Watch: The National Longitudinal Surveys.'' \emph{The Journal of Economic Perspectives} 15 (2): 239--53.

\leavevmode\hypertarget{ref-R}{}%
R Core Team. 2020. \emph{R: A Language and Environment for Statistical Computing}. Vienna, Austria: R Foundation for Statistical Computing. \url{https://www.R-project.org/}.

\leavevmode\hypertarget{ref-SingerJudithD2003Alda}{}%
Singer, Judith D, and John B Willett. 2003. \emph{Applied Longitudinal Data Analysis: Modeling Change and Event Occurrence}. Oxford u.a: Oxford Univ. Pr.

\leavevmode\hypertarget{ref-notiris}{}%
Stodel, Megan. 2020. ``Stop Using Iris.'' \url{https://www.meganstodel.com/posts/no-to-iris/}.

\leavevmode\hypertarget{ref-brolgar}{}%
Tierney, Nicholas, Di Cook, and Tania Prvan. 2020. \emph{brolgar: BRowse Over Longitudinal data Graphically and Analytically in R}. \url{https://github.com/njtierney/brolgar}.

\leavevmode\hypertarget{ref-tukey}{}%
Tukey, John W. (John Wilder). 1977. \emph{Exploratory Data Analysis}. Addison-Wesley Series in Behavioral Science. Reading, Mass.: Addison-Wesley Pub. Co.

\leavevmode\hypertarget{ref-rlm}{}%
UCLA: Statistical Consulting Group. 2021. ``Robust Regression \textbar{} R Data Analysis Examples.'' February 2021. \url{https://stats.idre.ucla.edu/r/dae/robust-regression/}.

\leavevmode\hypertarget{ref-LooMarkvander2018Sdcw}{}%
van der Loo, Mark, and Edwin de Jonge. 2018. \emph{Statistical Data Cleaning with Applications in R}.

\leavevmode\hypertarget{ref-validate}{}%
van der Loo, Mark P. J., and Edwin de Jonge. 2021. ``Data Validation Infrastructure for R.'' \emph{Journal of Statistical Software} 97 (10): 1--31. \url{https://doi.org/10.18637/jss.v097.i10}.

\leavevmode\hypertarget{ref-mass}{}%
Venables, W. N., and B. D. Ripley. 2002. \emph{Modern Applied Statistics with S}. Fourth. New York: Springer. \url{http://www.stats.ox.ac.uk/pub/MASS4}.

\leavevmode\hypertarget{ref-tsibble}{}%
Wang, Earo, Dianne Cook, and Rob J Hyndman. 2020. ``A New Tidy Data Structure to Support Exploration and Modeling of Temporal Data.'' \emph{Journal of Computational and Graphical Statistics} 29 (3): 466--78. \url{https://doi.org/10.1080/10618600.2019.1695624}.

\leavevmode\hypertarget{ref-plyr}{}%
Wickham, Hadley. 2011. ``The Split-Apply-Combine Strategy for Data Analysis.'' \emph{Journal of Statistical Software, Articles} 40 (1): 1--29. \url{https://doi.org/10.18637/jss.v040.i01}.

\leavevmode\hypertarget{ref-WickhamHadley2014TD}{}%
---------. 2014. ``Tidy Data.'' \emph{Journal of Statistical Software} 59 (10): 1--23.

\leavevmode\hypertarget{ref-ggplot2}{}%
---------. 2016. \emph{ggplot2: Elegant Graphics for Data Analysis}. Springer-Verlag New York. \url{https://ggplot2.tidyverse.org}.

\leavevmode\hypertarget{ref-stringr}{}%
---------. 2019. \emph{stringr: Simple, Consistent Wrappers for Common String Operations}. \url{https://CRAN.R-project.org/package=stringr}.

\leavevmode\hypertarget{ref-tidyr}{}%
---------. 2020. \emph{tidyr: Tidy Messy Data}. \url{https://CRAN.R-project.org/package=tidyr}.

\leavevmode\hypertarget{ref-tidyverse}{}%
Wickham, Hadley, Mara Averick, Jennifer Bryan, Winston Chang, Lucy D'Agostino McGowan, Romain François, Garrett Grolemund, et al. 2019. ``Welcome to the tidyverse.'' \emph{Journal of Open Source Software} 4 (43): 1686. \url{https://doi.org/10.21105/joss.01686}.

\leavevmode\hypertarget{ref-dplyr}{}%
Wickham, Hadley, Romain François, Lionel Henry, and Kirill Müller. 2020. \emph{dplyr: A Grammar of Data Manipulation}. \url{https://CRAN.R-project.org/package=dplyr}.

\leavevmode\hypertarget{ref-readr}{}%
Wickham, Hadley, and Jim Hester. 2020. \emph{readr: Read Rectangular Text Data}. \url{https://CRAN.R-project.org/package=readr}.

\leavevmode\hypertarget{ref-nlsy79edu}{}%
Wolpin, Kenneth I. 2005. ``National Longitudinal Survey of Youth 1979 Cohort, 1979-2016 (Rounds 1-28).'' Published by Bureau of Labor Statistics, U.S. Department of Labor. \url{https://www.bls.gov/opub/mlr/2005/02/art3full.pdf}.

\leavevmode\hypertarget{ref-knitr}{}%
Xie, Yihui. 2014. ``Knitr: A Comprehensive Tool for Reproducible Research in R.'' In \emph{Implementing Reproducible Computational Research}, edited by Victoria Stodden, Friedrich Leisch, and Roger D. Peng. Chapman; Hall/CRC. \url{http://www.crcpress.com/product/isbn/9781466561595}.

\leavevmode\hypertarget{ref-rmarkdown}{}%
Xie, Yihui, Christophe Dervieux, and Emily Riederer. 2020. \emph{R Markdown Cookbook}. Boca Raton, Florida: Chapman; Hall/CRC. \url{https://bookdown.org/yihui/rmarkdown-cookbook}.

\leavevmode\hypertarget{ref-kableExtra}{}%
Zhu, Hao. 2019. \emph{kableExtra: Construct Complex Table with 'kable' and Pipe Syntax}. \url{https://CRAN.R-project.org/package=kableExtra}.

>>>>

% tightlist command for lists without linebreak
\providecommand{\tightlist}{%
  \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}

% From pandoc table feature
\usepackage{longtable,booktabs,array}
\usepackage{calc} % for calculating minipage widths
% Correct order of tables after \paragraph or \subparagraph
\usepackage{etoolbox}
\makeatletter
\patchcmd\longtable{\par}{\if@noskipsec\mbox{}\fi\par}{}{}
\makeatother
% Allow footnotes in longtable head/foot
\IfFileExists{footnotehyper.sty}{\usepackage{footnotehyper}}{\usepackage{footnote}}
\makesavenoteenv{longtable}

% Pandoc citation processing
\newlength{\cslhangindent}
\setlength{\cslhangindent}{1.5em}
\newlength{\csllabelwidth}
\setlength{\csllabelwidth}{3em}
\newlength{\cslentryspacingunit} % times entry-spacing
\setlength{\cslentryspacingunit}{\parskip}
% for Pandoc 2.8 to 2.10.1
\newenvironment{cslreferences}%
  {}%
  {\par}
% For Pandoc 2.11+
\newenvironment{CSLReferences}[2] % #1 hanging-ident, #2 entry spacing
 {% don't indent paragraphs
  \setlength{\parindent}{0pt}
  % turn on hanging indent if param 1 is 1
  \ifodd #1
  \let\oldpar\par
  \def\par{\hangindent=\cslhangindent\oldpar}
  \fi
  % set entry spacing
  \setlength{\parskip}{#2\cslentryspacingunit}
 }%
 {}
\usepackage{calc}
\newcommand{\CSLBlock}[1]{#1\hfill\break}
\newcommand{\CSLLeftMargin}[1]{\parbox[t]{\csllabelwidth}{#1}}
\newcommand{\CSLRightInline}[1]{\parbox[t]{\linewidth - \csllabelwidth}{#1}\break}
\newcommand{\CSLIndent}[1]{\hspace{\cslhangindent}#1}

\usepackage{tcolorbox}
\usepackage{fontawesome}
\usepackage{color, colortbl}
\usepackage{booktabs}
\usepackage{longtable}
\usepackage{array}
\usepackage{multirow}
\usepackage{wrapfig}
\usepackage{float}
\usepackage{colortbl}
\usepackage{pdflscape}
\usepackage{tabu}
\usepackage{threeparttable}
\usepackage{threeparttablex}
\usepackage[normalem]{ulem}
\usepackage{makecell}
\usepackage{xcolor}
\begin{document}
\maketitle

@@

Our approach to cleaning builds heavily on the \texttt{tidyverse} approach (Wickham et al. 2019). The data is first organized into ``tidy data'' (Wickham 2014) and then further wrangled using a step-wise piping with split-apply-combine strategy for mutating new variables (Wickham 2011). Tidy data shouldn't be confused with ``tame data'' which Kim, Ismay, and Chunn (2018) coined to refer to textbook data sets suitable for teaching, particularly teaching statistics. The resulting (tame) data is provided in a new R package called \texttt{[CENSORED]} which includes the code so that the process is reproducible, and could be used to further refresh the data as new records are made available in the NLSY79 database.

@@

\begin{itemize}
\tightlist
\item
  \texttt{NLSY79.csv}: comma separated value format of the response data,
\item
  \texttt{NLSY79.dat}: alternative text format of the response data,
\item
  \texttt{NLSY79.NLSY79}: tagset of variables that can be uploaded to the website to recreate the data set, and
\item
  \texttt{NLSY79.R}: R script for reading the data into R and converting the variables' names and label into something more sensible.
\end{itemize}

@@

Thus we re-arrange and wrangle the data into tidy data form, columns corresponding to individual ID, year, job number, wage in dollars and the demographic variables. This is done using the \texttt{tidyverse} suite of packages (Wickham et al. 2019): \texttt{tidyr} (Wickham 2020) to pivot the data into long form, with \texttt{dplyr} (Wickham et al. 2020) and \texttt{stringr} (Wickham 2019) to mutate new variables from the downloaded data from the database, and code levels of factors by text wrangling. The long form of the data makes it possible to do these data transformations efficiently, and it is an intermediate step towards the final target data. The code for tidying the data are demonstrated at \url{CENSORED/articles/raw-to-input-data.html} but also described in the subsequent subsections.

@@

{\centering \includegraphics[width=0.85\linewidth]{/Users/cookd/numbats/yowie-paper/paper/figures/flowchart_blind}

@@

\hypertarget{refs}{}
\begin{CSLReferences}{1}{0}
\leavevmode\vadjust pre{\hypertarget{ref-iris-data}{}}%
Andereson, Edgar. 1935. {``The Irises of the Gaspe Peninsula.''} \emph{Bulletin of the American Iris Society} 59: 2--5.

\leavevmode\vadjust pre{\hypertarget{ref-nlsy79}{}}%
Bureau of Labor Statistics, U.S. Department of Labor. 2021a. {``National Longitudinal Survey of Youth 1979 Cohort, 1979-2016 (Rounds 1-28).''} Produced and distributed by the Center for Human Resource Research (CHRR), The Ohio State University. Columbus, OH, through \url{https://www.nlsinfo.org/bibliography-citing-nls-data}.

\leavevmode\vadjust pre{\hypertarget{ref-nlsy79guide}{}}%
---------. 2021b. {``National Longitudinal Survey of Youth 1979 Cohort, Topical Guide to the Data.''} \url{https://www.nlsinfo.org/content/cohorts/nlsy79/topical-guide/employment/work-experience}.

\leavevmode\vadjust pre{\hypertarget{ref-shiny}{}}%
Chang, Winston, Joe Cheng, JJ Allaire, Yihui Xie, and Jonathan McPherson. 2020. \emph{{shiny: Web Application Framework for R}}. \url{https://CRAN.R-project.org/package=shiny}.

\leavevmode\vadjust pre{\hypertarget{ref-Chatfield1985TIEo}{}}%
Chatfield, C. 1985. {``The Initial Examination of Data.''} \emph{Journal of the Royal Statistical Society. Series A. General} 148 (3): 214--53.

\leavevmode\vadjust pre{\hypertarget{ref-eliznlsy}{}}%
Cooksey, Elizabeth C. 2017. {``Using the National Longitudinal Surveys of Youth (NLSY) to Conduct Life Course Analyses.''} In \emph{Handbook of Life Course Health Development}, edited by Richard M. Lerner Neal Halfon Christoper B. Forrest, 561--77. Cham: Springer. https://doi.org/\url{https://doi.org/10.1007/978-3-319-47143-3_23}.

\leavevmode\vadjust pre{\hypertarget{ref-DasuTamraparni2003Edma}{}}%
Dasu, Tamraparni, and Theodore Johnson. 2003. \emph{Exploratory Data Mining and Data Cleaning}. Wiley Series in Probability and Statistics. Hoboken: WILEY.

\leavevmode\vadjust pre{\hypertarget{ref-janitor}{}}%
Firke, Sam. 2020. \emph{{janitor: Simple Tools for Examining and Cleaning Dirty Data}}. \url{https://CRAN.R-project.org/package=janitor}.

\leavevmode\vadjust pre{\hypertarget{ref-racismnotrace}{}}%
Fullilove, M. T. 1998. {``Comment: Abandoning ``Race" as a Variable in Public Health Research--an Idea Whose Time Has Come.''} \emph{American Journal of Public Health} 88 (9): 1297--98.

\leavevmode\vadjust pre{\hypertarget{ref-grimshaw}{}}%
Grimshaw, Scott D. 2015. {``A Framework for Infusing Authentic Data Experiences Within Statistics Courses.''} \emph{The American Statistician} 69 (4): 307--14. \url{https://doi.org/10.1080/00031305.2015.1081106}.

\leavevmode\vadjust pre{\hypertarget{ref-purrr}{}}%
Henry, Lionel, and Hadley Wickham. 2020. \emph{{purrr: Functional Programming Tools}}. \url{https://CRAN.R-project.org/package=purrr}.

\leavevmode\vadjust pre{\hypertarget{ref-penguins-data}{}}%
Horst, Allison Marie, Alison Presmanes Hill, and Kristen B Gorman. 2020. \emph{Palmerpenguins: Palmer Archipelago (Antarctica) Penguin Data}. \url{https://doi.org/10.5281/zenodo.3960218}.

\leavevmode\vadjust pre{\hypertarget{ref-HuebnerMariannePhD2016Asat}{}}%
Huebner, Marianne, Werner Vach, and Saskia le Cessie. 2016. {``A Systematic Approach to Initial Data Analysis Is Good Research Practice.''} \emph{The Journal of Thoracic and Cardiovascular Surgery} 151 (1): 25--27.

\leavevmode\vadjust pre{\hypertarget{ref-ilk2004}{}}%
Ilk, Ozlem. 2004. {``Exploratory Multivariate Longitudinal Data Analysis and Models for Multivariate Longitudinal Binary Data.''} PhD thesis, Iowa State University. \url{https://doi.org/10.31274/rtd-180813-11012}.

\leavevmode\vadjust pre{\hypertarget{ref-tamedata}{}}%
Kim, A. Y, C. Ismay, and J. Chunn. 2018. {``The Fivethirtyeight {R} Package: ``Tame Data" Principles for Introductory Statistics and Data Science Courses.''} \emph{Technology Innovations in Statistics Education} 11 (1). \url{https://doi.org/10.5070/T511103589}.

\leavevmode\vadjust pre{\hypertarget{ref-KollerManuel2016rARP}{}}%
Koller, Manuel. 2016. {``{robustlmm: An R Package for Robust Estimation of Linear Mixed-Effects Models}.''} \emph{Journal of Statistical Software} 75 (6): 1--24.

\leavevmode\vadjust pre{\hypertarget{ref-notaverage}{}}%
Moncrief, Marc. 2015. {``By the Numbers - the Average Australian Doesn't Exist ... Not a Single One of Us Is 'Normal'.''} \url{https://bit.ly/smh-not-normal}.

\leavevmode\vadjust pre{\hypertarget{ref-opendata}{}}%
Open Knowledge Foundation. 2021. {``Open Definition. Defining Open in Open Data, Open Content, and Open Knowledge.''} 2021. \url{http://opendefinition.org/od/2.1/en/}.

\leavevmode\vadjust pre{\hypertarget{ref-patchwork}{}}%
Pedersen, Thomas Lin. 2020. \emph{{patchwork: The Composer of Plots}}. \url{https://CRAN.R-project.org/package=patchwork}.

\leavevmode\vadjust pre{\hypertarget{ref-MichaelRPergamit2001DWTN}{}}%
Pergamit, Michael R., Charles R. Pierret, Donna S. Rothstein, and Jonathan R. Veum. 2001. {``Data Watch: The National Longitudinal Surveys.''} \emph{The Journal of Economic Perspectives} 15 (2): 239--53.

\leavevmode\vadjust pre{\hypertarget{ref-R}{}}%
R Core Team. 2020. \emph{R: A Language and Environment for Statistical Computing}. Vienna, Austria: R Foundation for Statistical Computing. \url{https://www.R-project.org/}.

\leavevmode\vadjust pre{\hypertarget{ref-SingerJudithD2003Alda}{}}%
Singer, Judith D, and John B Willett. 2003. \emph{Applied Longitudinal Data Analysis: Modeling Change and Event Occurrence}. Oxford u.a: Oxford Univ. Pr.

\leavevmode\vadjust pre{\hypertarget{ref-notiris}{}}%
Stodel, Megan. 2020. {``Stop Using Iris.''} \url{https://www.meganstodel.com/posts/no-to-iris/}.

\leavevmode\vadjust pre{\hypertarget{ref-brolgar}{}}%
Tierney, Nicholas, Di Cook, and Tania Prvan. 2020. \emph{{brolgar: BRowse Over Longitudinal data Graphically and Analytically in R}}. \url{https://github.com/njtierney/brolgar}.

\leavevmode\vadjust pre{\hypertarget{ref-tukey}{}}%
Tukey, John W. (John Wilder). 1977. \emph{Exploratory Data Analysis}. Addison-Wesley Series in Behavioral Science. Reading, Mass.: Addison-Wesley Pub. Co.

\leavevmode\vadjust pre{\hypertarget{ref-rlm}{}}%
UCLA: Statistical Consulting Group. 2021. {``{Robust Regression \textbar{} R Data Analysis Examples}.''} February 2021. \url{https://stats.idre.ucla.edu/r/dae/robust-regression/}.

\leavevmode\vadjust pre{\hypertarget{ref-validate}{}}%
van der Loo, Mark P. J., and Edwin de Jonge. 2021. {``Data Validation Infrastructure for {R}.''} \emph{Journal of Statistical Software} 97 (10): 1--31. \url{https://doi.org/10.18637/jss.v097.i10}.

\leavevmode\vadjust pre{\hypertarget{ref-LooMarkvander2018Sdcw}{}}%
van der Loo, Mark, and Edwin de Jonge. 2018. \emph{Statistical Data Cleaning with Applications in r}.

\leavevmode\vadjust pre{\hypertarget{ref-mass}{}}%
Venables, W. N., and B. D. Ripley. 2002. \emph{Modern Applied Statistics with s}. Fourth. New York: Springer. \url{http://www.stats.ox.ac.uk/pub/MASS4}.

\leavevmode\vadjust pre{\hypertarget{ref-tsibble}{}}%
Wang, Earo, Dianne Cook, and Rob J Hyndman. 2020. {``A New Tidy Data Structure to Support Exploration and Modeling of Temporal Data.''} \emph{Journal of Computational and Graphical Statistics} 29 (3): 466--78. \url{https://doi.org/10.1080/10618600.2019.1695624}.

\leavevmode\vadjust pre{\hypertarget{ref-plyr}{}}%
Wickham, Hadley. 2011. {``The Split-Apply-Combine Strategy for Data Analysis.''} \emph{Journal of Statistical Software, Articles} 40 (1): 1--29. \url{https://doi.org/10.18637/jss.v040.i01}.

\leavevmode\vadjust pre{\hypertarget{ref-WickhamHadley2014TD}{}}%
---------. 2014. {``Tidy Data.''} \emph{Journal of Statistical Software} 59 (10): 1--23.

\leavevmode\vadjust pre{\hypertarget{ref-ggplot2}{}}%
---------. 2016. \emph{{ggplot2: Elegant Graphics for Data Analysis}}. Springer-Verlag New York. \url{https://ggplot2.tidyverse.org}.

\leavevmode\vadjust pre{\hypertarget{ref-stringr}{}}%
---------. 2019. \emph{{stringr: Simple, Consistent Wrappers for Common String Operations}}. \url{https://CRAN.R-project.org/package=stringr}.

\leavevmode\vadjust pre{\hypertarget{ref-tidyr}{}}%
---------. 2020. \emph{{tidyr: Tidy Messy Data}}. \url{https://CRAN.R-project.org/package=tidyr}.

\leavevmode\vadjust pre{\hypertarget{ref-tidyverse}{}}%
Wickham, Hadley, Mara Averick, Jennifer Bryan, Winston Chang, Lucy D'Agostino McGowan, Romain François, Garrett Grolemund, et al. 2019. {``Welcome to the {tidyverse}.''} \emph{Journal of Open Source Software} 4 (43): 1686. \url{https://doi.org/10.21105/joss.01686}.

\leavevmode\vadjust pre{\hypertarget{ref-dplyr}{}}%
Wickham, Hadley, Romain François, Lionel Henry, and Kirill Müller. 2020. \emph{{dplyr: A Grammar of Data Manipulation}}. \url{https://CRAN.R-project.org/package=dplyr}.

\leavevmode\vadjust pre{\hypertarget{ref-readr}{}}%
Wickham, Hadley, and Jim Hester. 2020. \emph{{readr: Read Rectangular Text Data}}. \url{https://CRAN.R-project.org/package=readr}.

\leavevmode\vadjust pre{\hypertarget{ref-nlsy79edu}{}}%
Wolpin, Kenneth I. 2005. {``National Longitudinal Survey of Youth 1979 Cohort, 1979-2016 (Rounds 1-28).''} Published by Bureau of Labor Statistics, U.S. Department of Labor. \url{https://www.bls.gov/opub/mlr/2005/02/art3full.pdf}.

\leavevmode\vadjust pre{\hypertarget{ref-knitr}{}}%
Xie, Yihui. 2014. {``Knitr: A Comprehensive Tool for Reproducible Research in {R}.''} In \emph{Implementing Reproducible Computational Research}, edited by Victoria Stodden, Friedrich Leisch, and Roger D. Peng. Chapman; Hall/CRC. \url{http://www.crcpress.com/product/isbn/9781466561595}.

\leavevmode\vadjust pre{\hypertarget{ref-rmarkdown}{}}%
Xie, Yihui, Christophe Dervieux, and Emily Riederer. 2020. \emph{R Markdown Cookbook}. Boca Raton, Florida: Chapman; Hall/CRC. \url{https://bookdown.org/yihui/rmarkdown-cookbook}.

\leavevmode\vadjust pre{\hypertarget{ref-kableExtra}{}}%
Zhu, Hao. 2019. \emph{{kableExtra: Construct Complex Table with 'kable' and Pipe Syntax}}. \url{https://CRAN.R-project.org/package=kableExtra}.

\end{CSLReferences}
